Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-09-11 paper.qmd
Updated Figure 17

error_plot_mnist <- error_df |>
  ggplot(aes(x = tSNE1,
             y = tSNE2, 
             color = type, 
             group = ID)) +
  geom_point(alpha=0.5, 
             size = 0.1) +
  scale_colour_brewer(
    type = "qual",
    palette = 2
  ) + 
  theme(
      legend.position = "bottom",
      legend.text = element_text(size = 10),          # Increase legend text size
      legend.key.size = unit(1, "lines")              # Increase size of legend keys
  ) +
  guides(
      color = guide_legend(
          override.aes = list(size = 2.5, alpha = 1)      # Increase point size and opacity in legend
      )
  ) 
```

@@

```{r}
#| label: img-with-error

@@

img_error_within <- c(5173, 6671, 2964, 2464, 6319, 7388, 4303, 7650)

pixels_gathered_within <-  mnist_data |>
  mutate(instance = row_number()) |>
  gather(pixel, value, -Label, -instance) |>
  extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
  filter(instance %in% img_error_within)

imge_error_sample_within <- pixels_gathered_within |>
  ggplot(aes(x, y, fill = value)) +
  geom_tile() +
  facet_wrap(~ instance, ncol = 4) +
  coord_fixed() +
  scale_fill_continuous_sequential(palette = "Grays") +
  theme(strip.background = element_blank(),
        strip.text.x = element_blank(),
        legend.position = "none")

img_error_outside <- c(3554, 501, 4422, 6860, 3869, 1529, 4916, 6144)

@@

```{r}
#| label: fig-model-error-mnist
#| fig-cap: "(a) The $10\\text{-}D$ model error for tSNE layout in $\\gD{}$ of the MNIST digit 1 dataset. Most low model errors are distributed along the big non-linear clutser, while most large model errors are concentrated along the small cluster. (b) There are certain images that exhibit large errors due to their deviation from the non-linear cluster, which makes them anomalies. The images associated with large model errors shows different patterns of digit 1."
#| fig-pos: H
#| fig-width: 4
#| fig-height: 8
#| out-height: 80%

#imge_error_sample_within

error_plot_mnist / (imge_error_sample_outside) +
  plot_annotation(tag_levels = 'a') &
  theme(plot.tag.position  = c(0.92, 0.92),
        plot.tag = element_text(color = "grey70"))
```

>>>>

highd_error_2d_mnist <- error_df |>
  ggplot(aes(x = tSNE1,
             y = tSNE2, 
             color = type, 
             group = ID)) +
  geom_point(alpha=0.5, 
             size = 0.1) +
  scale_colour_brewer(
    type = "qual",
    palette = 2
  ) + 
  theme(
      legend.position = "bottom",
      legend.text = element_text(size = 10),          # Increase legend text size
      legend.key.size = unit(1, "lines")              # Increase size of legend keys
  ) +
  guides(
      color = guide_legend(
          override.aes = list(size = 2.5, alpha = 1)      # Increase point size and opacity in legend
      )
  ) 
```

@@

```{r}
#| label: img-diff-pos

@@

img_right_top <- c(6561, 454, 6397)

pixels_gathered_within <-  mnist_data |>
  mutate(instance = row_number()) |>
  gather(pixel, value, -Label, -instance) |>
  extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
  filter(instance %in% img_right_top)

right_top_img <- pixels_gathered_within |>
  ggplot(aes(x, y, fill = value)) +
  geom_tile() +
  facet_wrap(~ instance, ncol = 4) +
  coord_fixed() +
  scale_fill_continuous_sequential(palette = "Grays") +
  theme(strip.background = element_blank(),
        strip.text.x = element_blank(),
        legend.position = "none")

img_middle <- c(3443, 7404, 7242)

pixels_gathered_within <-  mnist_data |>
  mutate(instance = row_number()) |>
  gather(pixel, value, -Label, -instance) |>
  extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
  filter(instance %in% img_middle)

middle_img <- pixels_gathered_within |>
  ggplot(aes(x, y, fill = value)) +
  geom_tile() +
  facet_wrap(~ instance, ncol = 4) +
  coord_fixed() +
  scale_fill_continuous_sequential(palette = "Grays") +
  theme(strip.background = element_blank(),
        strip.text.x = element_blank(),
        legend.position = "none")

img_right_bottom <- c(7098, 6456, 2)

pixels_gathered_within <-  mnist_data |>
  mutate(instance = row_number()) |>
  gather(pixel, value, -Label, -instance) |>
  extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
  filter(instance %in% img_right_bottom)

right_bottom_img <- pixels_gathered_within |>
  ggplot(aes(x, y, fill = value)) +
  geom_tile() +
  facet_wrap(~ instance, ncol = 4) +
  coord_fixed() +
  scale_fill_continuous_sequential(palette = "Grays") +
  theme(strip.background = element_blank(),
        strip.text.x = element_blank(),
        legend.position = "none")

#img_error_outside <- c(3554, 501, 4422, 6860, 3869, 1529, 4916, 6144)
img_error_outside <- c(501, 1529, 6144)

@@

```{r}
#| label: fig-model-error-mnist
#| fig-cap: "(a) The $10\\text{-}D$ model error for tSNE layout in $\\gD{}$ of the MNIST digit 1 dataset. Most low model errors are distributed along the big non-linear clutser, while most large model errors are concentrated along the small cluster. (b) There are certain images that exhibit large errors due to their deviation from the non-linear cluster, which makes them anomalies. The images associated with large model errors shows different patterns of digit 1."
#| fig-pos: H
#| out-height: 80%

highd_error_2d_mnist + (imge_error_sample_outside / right_top_img / middle_img / right_bottom_img)
```
